---
title: "Figure 6"
author: "Alice MacQueen"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(bigsnpr)
library(here)
```

#

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")

datadir <- here("data")
site_v <- list(TX2 = "TX2", MO = "MO",  MI = "MI", 
               three_sites = c("TX2", "MO", "MI"))
phe_v <- c("PAN_LEN", "PRIM_BN", "SEC_BN")

subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Atlantic = "Atlantic", 
                 three_subpops = c("Atlantic", "Gulf", "Midwest"))

h2_table <- tibble()
prefix <- "pvdiv_panicles_2019_BLUPs_kinship_geno_subset_"

outputdir <- here("analysis", "heritability")
```

# Violin plots of diversity phenotypes at three sites, maybe with ecotype means shown

```{r}
phe <- readRDS(here("data/panicle_phenotypes_genotype_subset.rds"))
phe <- phe %>%
  mutate(phe_name = case_when(PHE == "PAN_LEN" ~ "PL",
                              PHE == "PRIM_BN" ~ "PBN",
                              PHE == "SEC_BN" ~ "SBN"))
phe$ECOTYPE_NNET <- factor(phe$ECOTYPE_NNET, levels = c("Upland", "Coastal", "Lowland", "Unknown"))
phe$SITE <- factor(phe$SITE, levels = c("PKLE", "CLMB", "KBSM"))
phe$phe_name <- factor(phe$phe_name, levels = c("PL", "PBN", "SBN"))
                  
                  
```

```{r}
ecotype_means <- phe %>%
  filter(ECOTYPE_NNET != "Unknown") %>%
  group_by(phe_name, SITE, ECOTYPE_NNET) %>%
  summarise(Ecotype_mean = mean(MEAS),
            Ecotype_se = sd(MEAS)/sqrt(n()/3), # sample size for SE should be number of unique genotypes, three panicle phenotype replicates per genotype, on average
            Ecotype_n = n()/3)
subpop_means <- phe %>%
  filter(SUBPOP != "4X") %>%
  group_by(phe_name, SITE, SUBPOP) %>%
  summarise(Subpop_mean = mean(MEAS),
            Subpop_se = sd(MEAS)/sqrt(n()/3), # sample size for SE should be number of unique genotypes, three panicle phenotype replicates per genotype, on average
            Subpop_n = n()/3)


panel_a <- phe %>%
  ggplot(aes(x = SITE)) + 
  facet_grid(rows = vars(phe_name), scales = "free_y") +
  geom_violin(aes(y = MEAS)) +
  geom_point(data = ecotype_means, mapping = aes(x = SITE, y = Ecotype_mean,
                                                 color = ECOTYPE_NNET)) +
  geom_errorbar(data = ecotype_means, aes(x = SITE, ymin = Ecotype_mean - Ecotype_se*2, 
                    ymax = Ecotype_mean + Ecotype_se*2,
                    color = ECOTYPE_NNET), width = 0.25) +
  scale_color_manual(values = c("#3366CC", "#FFCC00", "#CC0000"), 
                     name = "Ecotype") +
  theme(legend.position = "right") + 
  labs(x = "Site", y = "Phenotypic measurement")

panel_supp1 <- phe %>%
  ggplot(aes(x = SITE)) + 
  facet_grid(rows = vars(phe_name), scales = "free_y") +
  geom_violin(aes(y = MEAS)) +
  geom_point(data = subpop_means, mapping = aes(x = SITE, y = Subpop_mean,
                                                 color = SUBPOP)) +
  scale_color_manual(values = c("#6E91CB", "#F47F72", "#442C83")) +
  geom_errorbar(data = subpop_means, aes(x = SITE, ymin = Subpop_mean - Subpop_se*2, 
                    ymax = Subpop_mean + Subpop_se*2,
                    color = SUBPOP), width = 0.25) +
  theme(legend.position = "right") + 
  labs(x = "Site", y = "Phenotypic measurement")

phe %>%
  filter(SUBPOP != "4X" & !(SUBPOP == "Midwest" & ECOTYPE_NNET == "Coastal") &
            !(SUBPOP == "Atlantic" & ECOTYPE_NNET == "Lowland") & 
            !(SUBPOP == "Gulf" & ECOTYPE_NNET == "Upland")) %>%
  ggplot(aes(x = SUBPOP)) + 
  facet_grid(cols = vars(SITE), rows = vars(phe_name), scales = "free_y") +
  geom_violin(aes(y = MEAS, color = ECOTYPE_NNET)) +
  scale_color_manual(values = c("#3366CC", "#FFCC00", "#CC0000"))

panel_suppX1 <- phe %>%
  filter(SUBPOP != "4X" & !(SUBPOP == "Midwest" & ECOTYPE_NNET == "Coastal") &
            !(SUBPOP == "Atlantic" & ECOTYPE_NNET == "Lowland") & 
            !(SUBPOP == "Gulf" & ECOTYPE_NNET == "Upland")) %>%
  ggplot(aes(x = ECOTYPE_NNET)) + 
  facet_grid(cols = vars(SITE), rows = vars(phe_name), scales = "free_y") +
  geom_violin(aes(y = MEAS, color = SUBPOP)) +
  scale_color_manual(values = c("#6E91CB", "#F47F72", "#442C83"))
```

# heritability barplots

```{r}
h2_table <- read_csv(file.path(outputdir, paste0(prefix, "h2_df_no_GxE.csv")))

h2_table <- h2_table %>%
  mutate(SITE = case_when(site == "TX2" ~ "PKLE",
                          site == "MO" ~ "CLMB",
                          site == "MI" ~ "KBSM",
                          site == "three_sites" ~ "All"),
         phe_name = case_when(phe == "PAN_LEN" ~ "PL",
                              phe == "PRIM_BN" ~ "PBN",
                              phe == "SEC_BN" ~ "SBN")) 
h2_table$phe_name <- factor(h2_table$phe_name, levels = c("PL", "PBN", "SBN"))
h2_table$SITE <- factor(h2_table$SITE, levels = c("PKLE", "CLMB", "KBSM", "All"))

h2_specific <- filter(h2_table, SITE != "All" & subpop != "three_subpops")
h2_specific1 <- filter(h2_table, SITE != "All")
 
panel_b <- h2_table %>%
  filter(SITE != "All" & subpop == "three_subpops") %>%
  ggplot(aes(x = SITE, y = Estimate)) +
  geom_bar(aes(), stat = "identity", fill = "grey", color = "black") + 
  facet_grid(rows = vars(phe_name)) + #, scales = "free_y") +
  geom_errorbar(aes(ymin = Estimate - SE*2, 
                    ymax = Estimate + SE*2), width = 0.25) + 
  labs(x = "Site", y = bquote('h'^2))

panel_supp2 <- h2_specific1 %>%
  ggplot(aes(x = subpop, y = Estimate)) +
  facet_grid(cols = vars(SITE), rows = vars(phe_name)) +
  geom_point(aes(color = subpop)) +
  scale_color_manual(values = c("#6E91CB", "#F47F72", "#442C83", "grey")) +
  geom_errorbar(data = h2_specific1, aes(ymin = Estimate - SE*2, 
                    ymax = Estimate + SE*2, color = subpop), width = 0.25) +
  theme(legend.position = "right") +
  geom_hline(data = filter(h2_specific1, subpop == "three_subpops"), 
             mapping = aes(yintercept = Estimate + SE*2), color = "grey", 
             linetype = 2)  +
  geom_hline(data = filter(h2_specific1, subpop == "three_subpops"), 
             mapping = aes(yintercept = Estimate - SE*2), color = "grey", 
             linetype = 2) + 
  labs(x = "Site", y = bquote('h'^2))

h2_table %>%
  filter(SITE != "All" & subpop == "three_subpops") %>%
  arrange(phe)
```


```{r}
27*.05
.05/27
```

# Phenotypic correlations Fig 6c

```{r}
site_v2 <- list(PKLE = "PKLE", CLMB = "CLMB", KBSM = "KBSM")
site_plots <- list()

for (i in seq_along(site_v2)) {
  cov_df <- phe %>% filter(SITE %in% site_v2[[i]]) %>% 
    select(-PHE) %>%
    pivot_wider(names_from = "phe_name", values_from = "MEAS")
  cor_phe <- cor(cov_df[,(10:12)], use = "pairwise")

# Replace 1's on the diagonal with the coefficient of variation within each common garden.
#cov_sd <- matrixStats::colSds(as.matrix(cov_df[,10:ncol(cov_df)]), na.rm = TRUE)
# cov_mean <- matrixStats::colMeans2(as.matrix(cov_df[,10:ncol(cov_df)]), 
#                                   na.rm = TRUE)
# diag(cor_phe) <- cov_sd/cov_mean 

# If the correlation matrix is symmetric, remove the upper left half for plotting.
if(isSymmetric(cor_phe)){
      for(m in 1:nrow(cor_phe)){
        for(j in 1:ncol(cor_phe)){
          if(m < j){
            cor_phe[m, j] <- NA
          }
        }
      }
}

# Put matrix in long format for ggplot.
    U1 <- as_tibble(cor_phe, rownames = "rowU", .name_repair = "unique") %>%
            pivot_longer(cols = -.data$rowU, names_to = "colU",
                         values_to = "covar") %>%
            filter(!is.na(.data$covar))
    U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
    U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))

    site_plots[[i]] <- U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    switchgrassGWAS::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                  "white", "#27AD81FF", "#5DC863FF",
                                  "#FDE725FF"),
                         limits = c(-1, 1)) +
    #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    # Add text labels for each tile with the covariance fraction
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          #axis.text = element_blank(), # optionally remove this text
          axis.ticks = element_blank(),
          axis.line.x = element_blank(),
          axis.line.y = element_blank(),
          panel.spacing.x = unit(-0, 'cm')) +
    xlab("") + ylab("") + labs(title = names(site_v2)[i], 
                               fill = "Phenotypic\ncorrelation")
}


  cov_dfall <- phe %>%
    select(-PHE, -manu_site, -PLOT_GL) %>%
    pivot_wider(names_from = c("phe_name", "SITE"), values_from = "MEAS", values_fn = median)
  cor_phe <- cor(cov_dfall[,(7:15)], use = "pairwise")
```

# Mash results

```{r}
bigsnp_inputs <- read_delim(here("analysis", 
                                   "bigsnp_inputs.txt"),  delim = " ", 
                                   col_names = "SNPfiles") 
snp <- snp_attach(bigsnp_inputs$SNPfiles[7])
m <- readRDS(here("analysis", "mash",
          "Full_mash_model_5000_SNPs_U_ed_and_100000_SNPs_mash_fit_PLANT_ID_geno_subset.rds"))

library(snpdiver)
p_1 <- mash_plot_marker_effect(m, snp, n = 1)
p_2 <- mash_plot_marker_effect(m, snp, n = 3)
p_3 <- mash_plot_marker_effect(m, snp, n = 5)
p_4 <- mash_plot_marker_effect(m, snp, n = 7)
p_5 <- mash_plot_marker_effect(m, snp, n = 9)
p_6 <- mash_plot_marker_effect(m, snp, n = 1000)


p2 <- mash_plot_sig_by_condition(m)
p3 <- mash_plot_covar(m)
mash_plot_Ulist(m)
p4 <- mash_plot_manhattan_by_condition(m, snp, cond = c("PAN_LEN_TX2", "PAN_LEN_MO", "PAN_LEN_MI"))
p5 <- mash_plot_pairwise_sharing(m)
p6 <- mash_plot_Ulist(m)

gxe_df <- get_GxE(m, thr.r2 = 0)  # thr.r2 does not work yet
saveRDS(gxe_df, file = here("analysis/mash/GxE_df_full_mash_model_5000_SNPs_U_ed_and_100000_SNPs_mash_fit_PLANT_ID_geno_subset.rds"))
length(get_significant_results(m))
48271/18704908

35971/48271
sum(gxe_df$S_2_no)
sum(gxe_df$S_all_pairwise)
sum(gxe_df$S_DS)
sum(gxe_df$S_AP)
sum(gxe_df$S_1_row)
sum(gxe_df$S_CN)
sum(gxe_df$S_1_col)
77612+172846+3405271+567822
4050705-3405271-77612-567822
197488+172846

# all_pairwise - 2_no - AP - CN
sum(gxe_df$S_2_no)/sum(gxe_df$S_all_pairwise) # 84.07% No GxE
sum(gxe_df$S_AP)/sum(gxe_df$S_all_pairwise)   # 1.92% AP
sum(gxe_df$S_CN)/sum(gxe_df$S_all_pairwise)   # 14.02% CN
```

### Plot histogram of enrichments of random QTL intervals
```{r}
sig_random <- readRDS(here("analysis", "mash",
                     paste0("1000_Random_QTL_df_significant_enrichments_of", 
                            "_mash_BF_1per_tailclumped_PLANT_ID_geno_subset",
                            ".rds")))

panel_e <- tibble(`mash significant enrichments` = sig_random) %>%
  arrange(desc(`mash significant enrichments`)) %>%
  ggplot(aes(x = `mash significant enrichments`)) +
  switchgrassGWAS::theme_oeco +
  geom_histogram(binwidth = 1) + 
  xlim(c(1, 18)) + ylab("") +
  geom_vline(xintercept = 10, linetype = 2, color = "red")

```

##### ---------

# clumping with r2 of 0
```{r}
S <- snpdiver:::get_log10bf(m)

mash_clumps_0r2 <- snp_clumping(pavir_snp$genotypes, 
                                infos.chr = markers$CHRN, thr.r2 = 0, 
                                infos.pos = markers$POS, 
                                S = S)
  mash_df_clumped <- manhattan$ggman_df[mash_clumps,]
  write_csv(mash_df_clumped, file = file.path(outputdir,
                                              paste0("Clumped_mash_output_df_",
                                                     suffix, ".csv")))
```

Only 303 markers are present when markers are clumped using the log10BF and using an r^2 of 0.
```{r}
stdev <- readRDS(here("analysis/mash/Mash_PLANT_ID_geno_subset_unlinked_0r2_stddev.rds"))
markers <- readRDS(here("analysis/mash/Mash_PLANT_ID_geno_subset_unlinked_0r2_markers.rds"))
lfsr <- readRDS(here("analysis/mash/Mash_PLANT_ID_geno_subset_unlinked_0r2_lfsr_sig.rds"))
eff_means <- readRDS(here("analysis/mash/Mash_PLANT_ID_geno_subset_unlinked_0r2_effect_means.rds"))
snppos <- readRDS(here("analysis/mash/Mash_PLANT_ID_geno_subset_clumped_0r2_SNPfile_positions.rds"))
```

```{r}
markers %>%
  ggplot(aes(x = POS, y = log10BF)) +
  geom_point(aes(color = maf)) +
  facet_wrap(vars(CHR), scales = "free_x") +
  geom_hline(yintercept = 3, linetype = 2) +
  theme(panel.spacing.x = unit(0.01, 'cm'))

markers %>%
  filter(log10BF > 3)
100/(77/9)
100/(77/3)

markers 
eff_means
lfsr
colnames(lfsr)[2:10] <- paste0(colnames(lfsr)[2:10], "_lfsr")
stdev <- as_tibble(stdev)

markers_0r2 <- markers %>%
  add_column(PL_PKLE_Effect_Mean = eff_means$PAN_LEN_TX2_Effect_Mean,
             PBN_PKLE_Effect_Mean = eff_means$PRIM_BN_TX2_Effect_Mean,
             SBN_PKLE_Effect_Mean = eff_means$SEC_BN_TX2_Effect_Mean,
             PL_CLMB_Effect_Mean = eff_means$PAN_LEN_MO_Effect_Mean,
             PBN_CLMB_Effect_Mean = eff_means$PRIM_BN_MO_Effect_Mean,
             SBN_CLMB_Effect_Mean = eff_means$SEC_BN_MO_Effect_Mean,
             PL_KBSM_Effect_Mean = eff_means$PAN_LEN_MI_Effect_Mean,
             PBN_KBSM_Effect_Mean = eff_means$PRIM_BN_MI_Effect_Mean,
             SBN_KBSM_Effect_Mean = eff_means$SEC_BN_MI_Effect_Mean,
             PL_PKLE_lfsr = lfsr$PAN_LEN_TX2_lfsr,
             PBN_PKLE_lfsr = lfsr$PRIM_BN_TX2_lfsr,
             SBN_PKLE_lfsr = lfsr$SEC_BN_TX2_lfsr,
             PL_CLMB_lfsr = lfsr$PAN_LEN_MO_lfsr,
             PBN_CLMB_lfsr = lfsr$PRIM_BN_MO_lfsr,
             SBN_CLMB_lfsr = lfsr$SEC_BN_MO_lfsr,
             PL_KBSM_lfsr = lfsr$PAN_LEN_MI_lfsr,
             PBN_KBSM_lfsr = lfsr$PRIM_BN_MI_lfsr,
             SBN_KBSM_lfsr = lfsr$SEC_BN_MI_lfsr,
             PL_PKLE_Effect_SD = stdev$PAN_LEN_TX2_Effect_SD,
             PBN_PKLE_Effect_SD = stdev$PRIM_BN_TX2_Effect_SD,
             SBN_PKLE_Effect_SD = stdev$SEC_BN_TX2_Effect_SD,
             PL_CLMB_Effect_SD = stdev$PAN_LEN_MO_Effect_SD,
             PBN_CLMB_Effect_SD = stdev$PRIM_BN_MO_Effect_SD,
             SBN_CLMB_Effect_SD = stdev$SEC_BN_MO_Effect_SD,
             PL_KBSM_Effect_SD = stdev$PAN_LEN_MI_Effect_SD,
             PBN_KBSM_Effect_SD = stdev$PRIM_BN_MI_Effect_SD,
             SBN_KBSM_Effect_SD = stdev$SEC_BN_MI_Effect_SD)

markers_0r2 <- markers_0r2 %>%
  mutate(PL_PKLE_Var_Exp = 2*maf*(1-maf)*PL_PKLE_Effect_Mean^2,
         PBN_PKLE_Var_Exp = 2*maf*(1-maf)*PBN_PKLE_Effect_Mean^2,
         SBN_PKLE_Var_Exp = 2*maf*(1-maf)*SBN_PKLE_Effect_Mean^2,
         PL_CLMB_Var_Exp = 2*maf*(1-maf)*PL_CLMB_Effect_Mean^2,
         PBN_CLMB_Var_Exp = 2*maf*(1-maf)*PBN_CLMB_Effect_Mean^2,
         SBN_CLMB_Var_Exp = 2*maf*(1-maf)*SBN_CLMB_Effect_Mean^2,
         PL_KBSM_Var_Exp = 2*maf*(1-maf)*PL_KBSM_Effect_Mean^2,
         PBN_KBSM_Var_Exp = 2*maf*(1-maf)*PBN_KBSM_Effect_Mean^2,
         SBN_KBSM_Var_Exp = 2*maf*(1-maf)*SBN_KBSM_Effect_Mean^2
         ) 
```

MAF in the three different subpopulations for these top SNPs
```{r}
atlantic <- phe %>% filter(SUBPOP == "Atlantic") %>%
  select(PLANT_ID) %>% unique()
midwest <- phe %>% filter(SUBPOP == "Midwest") %>%
  select(PLANT_ID) %>% unique()
gulf <- phe %>% filter(SUBPOP == "Gulf") %>%
  select(PLANT_ID) %>% unique()
maf381 <- phe %>%
   select(PLANT_ID) %>% unique()

atlanticn <- which(snp$fam$sample.ID %in% atlantic$PLANT_ID)
midwestn <- which(snp$fam$sample.ID %in% midwest$PLANT_ID)
gulfn <- which(snp$fam$sample.ID %in% gulf$PLANT_ID)
maf381n <- which(snp$fam$sample.ID %in% maf381$PLANT_ID)

atlanticmaf <- snp_MAF(snp$genotypes, ind.row = atlanticn, ind.col = snppos)
midwestmaf <- snp_MAF(snp$genotypes, ind.row = midwestn, ind.col = snppos)
gulfmaf <- snp_MAF(snp$genotypes, ind.row = gulfn, ind.col = snppos)
maf381 <- snp_MAF(snp$genotypes, ind.row = maf381n, ind.col = snppos)

markers_0r2 <- markers_0r2 %>%
  add_column(MAF_Atlantic = atlanticmaf, MAF_Gulf = gulfmaf, MAF_Midwest = midwestmaf)
markers_0r2 <- markers_0r2 %>%
  add_column(MAF_382 = maf381)
```

```{r}
markers_0r2 %>%
  filter(log10BF > 1.3)

markers_0r2_longer <- markers_0r2 %>%
  pivot_longer(cols = 7:42, names_to = "Meas_type", values_to = "MEAS") %>%
  mutate(Meas_type = case_when(grepl("Effect_SD", Meas_type) ~ str_replace(Meas_type, "Effect_SD", "EffectSD"),
                               grepl("Effect_Mean", Meas_type) ~ str_replace(Meas_type, "Effect_Mean", "EffectMean"),
                               grepl("Var_Exp", Meas_type) ~ str_replace(Meas_type, "Var_Exp", "VarExp"),
                               TRUE ~ Meas_type)) %>%
  separate(Meas_type, into = c("PHE", "SITE", "Meas_type"), sep = "_") %>%
  pivot_wider(names_from = Meas_type, values_from = MEAS) %>%
  arrange(CHR, POS, PHE)
```

```{r}
markers_0r2_longer %>%
  filter(log10BF > 2) %>%
  group_by(PHE, SITE) %>%
  summarise(sum(VarExp))
markers_0r2_longer %>%
  filter(log10BF > 3) %>%
  group_by(PHE, SITE) %>%
  summarise(sum(VarExp))
markers_0r2_longer %>%
  mutate(VarExp_381 = MAF_382*(1-MAF_382)*EffectMean^2) %>%
  filter(lfsr < 0.05) %>%
  group_by(PHE, SITE) %>%
  summarise(VarExp_381 = sum(VarExp_381))

markers_0r2_longer %>%
  filter(lfsr < 0.05) %>%
  group_by(CHR, POS) %>%
  tally()
markers_0r2_longer %>%
  filter(log10BF > 1.3 & lfsr < 0.05) %>%
  group_by(CHR, POS) %>%
  tally()
markers_0r2_longer %>%
  filter(lfsr < 0.05) %>%
  group_by(CHR, POS) %>%
  tally() %>%
  group_by(n) %>%
  tally()

187/303
get_significant_results(m)[1:10]
```

```{r}
82/(82+11+22+23+19+15+13+1+1)

(105+77+77+63+60+88)/6*52
4073/2/2/2/2/2
2^12
(73-64)/52
(105+77+77+63+60+88)/6+(73-64)/52
78.77*52
35*52-15
4096-1806
```

```{r}
p_1
p_2
p_3
p_4
p_5
p_6
```


```{r}
markers_0r2_longer <- markers_0r2_longer %>%
  mutate(MAF_specific = case_when(MAF_Atlantic > MAF_382 & MAF_Gulf < MAF_382 & 
                                    MAF_Midwest < MAF_382 ~ "Atlantic",
                                  MAF_Atlantic < MAF_382 & MAF_Gulf > MAF_382 &
                                    MAF_Midwest < MAF_382 ~ "Gulf",
                                  MAF_Atlantic < MAF_382 & MAF_Gulf < MAF_382 & 
                                    MAF_Midwest > MAF_382 ~ "Midwest",
                                  TRUE ~ "Shared")) 

markers_0r2_longer %>%
  group_by(MAF_specific) %>%
  tally() %>%
  mutate(n = n/9)

```


```{r}
markers_0r2_longer$SITE <- factor(markers_0r2_longer$SITE, levels = c("PKLE", "CLMB", "KBSM"))
markers_0r2_longer$MAF_specific <- factor(markers_0r2_longer$MAF_specific, levels = c("Gulf", "Midwest", "Atlantic", "Shared"))
markers_0r2_longer$PHE <- factor(markers_0r2_longer$PHE, levels = c("PL", "PBN", "SBN"))

panel_f <- markers_0r2_longer %>%
  mutate(pos_label = paste(round(POS/1000000, 2), "Mb", sep = " ")) %>%
  filter(log10BF > 5) %>%
  ggplot(aes(x = SITE, y = EffectMean, color = MAF_382)) +
  geom_point(aes()) +
  facet_grid(rows = vars(PHE), cols = vars(CHR, pos_label)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_errorbar(aes(ymin = EffectMean - EffectSD, ymax = EffectMean + EffectSD)) +
  theme(panel.spacing.x = unit(0.005, 'cm'),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "right") +
  scale_color_viridis_c(option = "B", limits = c(0,0.1), name = "MAF") +
  labs(y = "SNP Effects", x = "Site")

markers_0r2_longer %>%
  filter(log10BF > 5) %>%
  ggplot(aes(x = SITE, y = EffectMean)) +
  geom_point() +
  facet_grid(rows = vars(CHR, POS), cols = vars(PHE)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_errorbar(aes(ymin = EffectMean - EffectSD, ymax = EffectMean + EffectSD)) +
  coord_flip() +
  theme(panel.spacing.x = unit(0.02, 'cm')) 

markers_0r2_longer %>%
  filter(log10BF > 2 & CHR %in% c("Chr09K", "Chr09N")) %>%
  ggplot(aes(x = SITE, y = EffectMean, color = MAF_specific)) +
  geom_point() +
  facet_grid(rows = vars(PHE), cols = vars(CHR, POS)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_errorbar(aes(ymin = EffectMean - EffectSD, ymax = EffectMean + EffectSD)) +
  theme(panel.spacing.x = unit(0.01, 'cm'),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  +
  scale_color_manual(values = c("#6E91CB", "#F47F72", "#442C83", "grey"))

saveRDS(markers_0r2_longer, file = here("analysis/mash/Marker_summary_statistics_long_mash_PLANT_ID_geno_subset_0r2.rds"))
```


# Figure 6

Put together Figure 6 from the panels built above
```{r}

```


```{r}
legend_corr <- get_legend(p6$simple_het_3_ggobject)


panel_c <- ggdraw() +
  draw_plot(site_plots[[1]], x = 0, y = 0, width = 0.3, height = 1) +
  draw_plot(site_plots[[2]], x = 0.3, y = 0, width = 0.3, height = 1) +
  draw_plot(site_plots[[3]], x = 0.6, y = 0, width = 0.3, height = 1) +
  draw_plot(legend_corr, x = 0.9, y = 0, width = 0.1, height = 1)

```

```{r}
library(ggrepel)
# install.packages("ggrepel")
gxe_df
panel_d <- p3$covar_df %>%
  filter(Mass > 0) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(stat = "identity") + 
  coord_flip() +
  geom_label_repel(aes(label = round(Mass, 5)), direction = "x", size = 3)
18700000*.00063725

markers_0r2_longer %>%
  mutate(VarExp_381 = MAF_382*(1-MAF_382)*EffectMean^2) %>%
  filter(lfsr < 0.05) %>%
  group_by(PHE, SITE) %>%
  summarise(VarExp_381 = sum(VarExp_381))
markers_0r2_longer %>%
  mutate(VarExp_381 = MAF_382*(1-MAF_382)*EffectMean^2) %>%
  filter(lfsr < 0.05) %>%
  group_by(PHE, SITE) %>%
  tally()
gxe_df
markers_0r2_longer %>%
  ggplot(aes(x = MAF_382)) +
  geom_histogram()
```

```{r}
36.3/83.93
17.445/75.22
19/80.21
79.21/84.14
84.5987/81.27
75/78.06
17.4/68.09
20.36/64.53
23.4/58.83


```


```{r}
panel_a   # phenotypic distributions
panel_b   # heritabilities
panel_c   # small phenotypic correlations by site - 
panel_d   # covar with > 0.01% posterior weight
panel_e   # mash and QTL overlap
panel_f

(figure6 <- ggdraw() +
  draw_plot(panel_a, x = 0, y = 0.5, width = 0.3, height = 0.5) +
  draw_plot(panel_b, x = 0.3, y = 0.5, width = 0.2, height = 0.5) +
  draw_plot(panel_c, x = 0.5, y = 0.75, width = 0.5, height = 0.25) +
  draw_plot(panel_d, x = 0.5, y = 0.5, width = 0.25, height = 0.25) +
  draw_plot(panel_e, x = 0.75, y = 0.5, width = 0.25, height = 0.25) +
  draw_plot(panel_f, x = 0, y = 0, width = 1, height = 0.5) +
  draw_label("a", x = 0.01, y = .98) +
  draw_label("b", x = 0.31, y = .98) +
  draw_label("c", x = 0.51, y = .98) +
  draw_label("d", x = 0.51, y = .76) +
  draw_label("e", x = 0.74, y = .76) +
  draw_label("f", x = 0.01, y = .49) +
  draw_label("MAF", x = 0.955, y = .37, size = 11) +
  draw_label("Correlation", x = 0.96, y = .98, size = 11) +
  draw_label("Ecotype", x = 0.255, y = .905, size = 11)
    
    ) 
  


save_plot(filename = here("analysis/mash/Figure6.svg"), figure6, base_height = 7, base_width = 13)


```


