---
title: "Data Setup"
author: "Alice MacQueen"
date: 2021-06-30
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(asreml)
library(snpdiver)
library(cowplot)
library(matrixStats)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# 

This is a genetic analysis of three panicle traits collected in the diversity panel in 2019 at three common gardens: KBS (MI), Columbia (MO), and Pickle (TX2).

The three traits are PAN_LEN, PRIM_BN, and SEC_BN: 

  - PAN_LEN: length from first primary branch to tip of panicle
  - PRIM_BN: # primary branches
  - SEC_BN: # secondary branches off the first primary branch.

These traits were also measured in the fourway cross, though at more gardens.

# Prepare phenotypes

```{r}
prefix = "pvdiv_panicles_2019_"
workingdir <- file.path("~", "Github", "PanicleData", "9-DiversityPanel")
datadir <- file.path("~", "Github", "pvdiv-phenotypes", "data")
outputdir <- file.path(workingdir, "data")
phenotypes <- read_rds(file.path(datadir, "Phenotypes_cleaned.rds"))
metadata <- readRDS(file.path(datadir, "metadata.rds"))
saveRDS(metadata, file = file.path(outputdir, "metadata.rds"))
```

```{r}
panicles <- phenotypes %>%
  filter(grepl("PAN_LEN", PHE) | grepl("PRIM_BN", PHE) | grepl("SEC_BN", PHE)) %>%
  pivot_wider(id_cols = c(PLANT_ID, SUBPOP), names_from = c("manu_site", "PHE"), 
              values_from = MEAS, values_fn = mean) %>%
  filter(SUBPOP !="8X")

pan_subset <- panicles %>%
  left_join(select(metadata, PLANT_ID, ECOTYPE_NNET)) %>%
  mutate(Any_NA = rowSums(!is.na(panicles))) %>%
  select(PLANT_ID, SUBPOP, ECOTYPE_NNET, Any_NA, everything()) %>%
  filter(Any_NA > 21) %>%
  arrange(Any_NA)

pan_subset_phe <- phenotypes %>%
  filter(PLANT_ID %in% pan_subset$PLANT_ID) %>%
  filter(grepl("PAN_LEN", PHE) | grepl("PRIM_BN", PHE) | grepl("SEC_BN", PHE)) %>%
  filter(SITE %in% c("KBSM", "CLMB", "PKLE")) %>%
  mutate(MEAS = case_when(MEAS == 105601 ~ 10,
                   TRUE ~ MEAS)) %>%
  separate(PHE, into = c("PHE", NA, "REP"), sep = c(-2,-1)) %>%
  left_join(select(metadata, PLANT_ID, ECOTYPE_NNET), by = "PLANT_ID") %>%
  select(PLANT_ID, SUBPOP, ECOTYPE_NNET, everything())

pan_phe <- phenotypes %>%
  #filter(PLANT_ID %in% pan_subset$PLANT_ID) %>%
  filter(grepl("PAN_LEN", PHE) | grepl("PRIM_BN", PHE) | grepl("SEC_BN", PHE)) %>%
  filter(SITE %in% c("KBSM", "CLMB", "PKLE")) %>%
  mutate(MEAS = case_when(MEAS == 105601 ~ 10,
                   TRUE ~ MEAS)) %>%
  separate(PHE, into = c("PHE", NA, "REP"), sep = c(-2,-1)) %>%
  left_join(select(metadata, PLANT_ID, ECOTYPE_NNET), by = "PLANT_ID") %>%
  select(PLANT_ID, SUBPOP, ECOTYPE_NNET, everything())

saveRDS(pan_subset_phe, file = file.path(outputdir, "panicle_phenotypes_genotype_subset.rds"))
saveRDS(pan_phe, file = file.path(outputdir, "panicle_phenotypes.rds"))
```

Full set vs subset of genotypes, what subpopulations and ecotypes are lost?
```{r}
panicles %>%
  left_join(select(metadata, PLANT_ID, ECOTYPE_NNET)) %>%
  mutate(Any_NA = rowSums(!is.na(panicles))) %>%
  select(PLANT_ID, SUBPOP, Any_NA, everything()) %>%
  filter(Any_NA > 21) %>%
  group_by(SUBPOP) %>% #, ECOTYPE_NNET) %>%
  tally()

panicles %>%
  group_by(SUBPOP) %>%
  tally()

panicles %>%
  left_join(select(metadata, PLANT_ID, ECOTYPE_NNET)) %>%
  mutate(Any_NA = rowSums(!is.na(panicles))) %>%
  select(PLANT_ID, SUBPOP, Any_NA, everything()) %>%
  filter(Any_NA > 21) %>%
  group_by(SUBPOP, ECOTYPE_NNET) %>%
  tally()

panicles %>%
  left_join(select(metadata, PLANT_ID, ECOTYPE_NNET)) %>%
  group_by(SUBPOP, ECOTYPE_NNET) %>%
  tally()
```


4X: 43 to 27  (62.7%)
Atlantic: 267 to 164  (61.4%)
Gulf: 228 to 129   (56.6%)
Midwest: 133 to 89  (66.9%)

60 vs 108 Atlantic Coastal (55.55%)
101 vs 148 Atlantic Upland (68.2%)
39 vs 85 Gulf Coastal (45.9%)
89 vs 133 Gulf Lowland (66.9%)
87 vs 125 Midwest Upland (69.6%)

# ---------

# BLUPs using PLANT_ID

There are three technical replicates measured for each PLANT_ID and each phenotype. Get the BLUP for each of these by just PLANT_ID.
```{r}
workingdir <- file.path("~", "Github", "PanicleData", "9-DiversityPanel")
datadir <- file.path(workingdir, "data")
analysisdir <- file.path(workingdir, "analysis")
site_v <- list(TX2 = "TX2", MO = "MO",  MI = "MI")
h2_table <- tibble()
phe_v <- c("PAN_LEN", "PRIM_BN", "SEC_BN")
```


```{r}
prefix <- "pvdiv_panicles_2019_BLUPs_PLANT_ID_geno_subset_"
phe <- readRDS(file.path(datadir, "panicle_phenotypes_genotype_subset.rds"))
```

```{r}
prefix <- "pvdiv_panicles_2019_BLUPs_PLANT_ID_"
phe <- readRDS(file.path(datadir, "panicle_phenotypes.rds"))
```


```{r}
i = 1 ; k = 1
outputdir <- file.path(workingdir, "analysis", "PLANT_ID_BLUPs")
if(!dir.exists(outputdir)){
  dir.create(outputdir)
}
if(!dir.exists(file.path(outputdir, "ASReml_plots"))){
  dir.create(file.path(outputdir, "ASReml_plots"))
}
if(!dir.exists(file.path(outputdir, "ASReml_BLUPs"))){
  dir.create(file.path(outputdir, "ASReml_BLUPs"))
}

for(i in seq_along(site_v)){ # may need to adjust h2_table line below 
  for(k in seq_along(phe_v)){
      phe_single <- phe %>%
        filter(manu_site %in% site_v[[i]] & PHE %in% phe_v[[k]]) %>%
        rename(phenotype = MEAS) %>%
        mutate(PLANT_ID = as_factor(PLANT_ID),
               SITE = as_factor(SITE),
               PLOT_GL = as_factor(PLOT_GL)) %>%
        as.data.frame()

  tryCatch({
  asr_out <- asreml(phenotype ~ 1,
                     random = ~idv(PLANT_ID),
                     residual = ~idv(units),
                     data = phe_single, 
                     workspace = "3gb")
  
  blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
    as_tibble(rownames = "Effect")
  write_csv(blup, file.path(outputdir, "ASReml_BLUPs", paste0(prefix,
                                              names(site_v)[i], "_",
                                              phe_v[k], ".csv")))
}, error = function(e){cat("ERROR :",conditionMessage(e), "\n")})

tryCatch({
    save_plot(filename = file.path(outputdir, "ASReml_plots", paste0(prefix,
                                                   "ASReml_performance_",
                                                   names(site_v)[i], "_",
                                                    phe_v[k], ".png")), 
          plot = plot(asr_out))
  }, error = function(e){cat("ERROR :",conditionMessage(e), "\n")})

    }
  }
```

## Read in BLUP csv files

```{r}
phe_gwas <- tibble()

    for(i in seq_along(site_v)){
      for(k in seq_along(phe_v)){
        phe_raw <- read_csv(file.path(outputdir, "ASReml_BLUPs",
                                      paste0(prefix,
                                              names(site_v)[i], "_",
                                              phe_v[k], ".csv")))
        names(phe_raw)[2] <- paste0(phe_v[k], "_", names(site_v)[i])
        phe_proc <- phe_raw %>%
          separate(Effect, into = c(NA, "PLANT_ID"), sep = 9) %>%
          select(PLANT_ID, 2)
        if(i == 1 & k == 1){
          phe_gwas <- phe_proc
        } else {
          phe_gwas <- phe_gwas %>%
            left_join(phe_proc, by = "PLANT_ID")
        }
      }
    }
  write_csv(phe_gwas, file = file.path(outputdir,
                                       paste0(prefix, "BLUP_phenotypes",
                                              ".csv")))
```

